---
title: "Engineering Resource Analysis and Profit Forecasting"
author: "Abdulfetah"
date: "2025-03-18"
output:
  html_document: default
  pdf_document: default
---

```{r setup_2, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
```

```{r setup, include=FALSE}

install.packages("forecast", repos = "https://cran.rstudio.com/")  
library(forecast) 
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(stringr)
library(tidyr)
library(rmarkdown)
library(knitr)

```

Summary
Timeless Transport Models operates three factories producing various product lines. The company is aiming to forecast profits for the second half of 2005. 
The questions to address are:

How many of each product line do they expect to sell monthly between June 2005 and December 2005?
What is the cost per product line to produce?
What is the average profit per product line?
Given the average profit and forecast sales per product, how much profit should the company expect in the second half of 2005?
The analysis uses historical data from 2003 to 2004 to forecast sales for the second half of 2005 using the ARIMA model and calculates the expected profit based on these fore.

Data Overview:
```{r}
production_data <- read_csv("C:/Users/abdu/Engineering Resource Analysis and Allocation/Engineering-Resource-Analysis-and-Allocation/data/weekly_factory_production.csv")

production_data$week <- as.Date(production_data$week, format="%m/%d/%Y")

# Check the first few rows
head(production_data)

```
Data Stats: it includes columns for week, factory_id, weekly_cost, and product quantities for Classic Cars, Motorcycles, Planes, Ships, Trucks and Buses, and Vintage Cars.
The data spans from January 2003 to April 2004.
Data has been aggregated by month to simplify analysis.
Aggregate Weekly Data to Monthly Data:

```{r}
# Aggregate data by month
production_data$month <- floor_date(production_data$week, "month")

monthly_data <- production_data %>%
  group_by(month) %>%
  summarise(
    total_cost = sum(weekly_cost),
    total_classic_cars = sum(`Classic Cars`),
    total_motorcycles = sum(Motorcycles),
    total_planes = sum(Planes),
    total_ships = sum(Ships),
    total_trucks_and_buses = sum(`Trucks and Buses`),
    total_vintage_cars = sum(`Vintage Cars`)
  )
# Display aggregated monthly data
head(monthly_data)


```
Analysis:
How Many of Each Product Line Do They Expect to Sell Monthly Between June 2005 and December 2005?
We used the ARIMA model to forecast sales for each product line from June to December 2005. Below is the forecast for Classic Cars.

```{r}

# Time series forecasting for Classic Cars
ts_classic <- ts(monthly_data$total_classic_cars, start = c(2003, 1), frequency = 12)
forecast_classic <- forecast(auto.arima(ts_classic), h = 7)

# Average profit per unit
selling_price_per_unit <- 500
cost_per_unit <- 300  # Add this line
avg_profit_per_unit <- selling_price_per_unit - cost_per_unit
avg_profit_per_unit

# Display forecasted values
forecast_classic_df <- data.frame(forecast_classic)
kable(forecast_classic_df, caption = "Classic Cars Monthly Forecast (Jun-Dec 2005)")
autoplot(forecast_classic) +
  labs(title = "Forecast of Classic Cars Production (Jun–Dec 2005)", y = "Units", x = "Month")


```
 What is the Cost Per Product Line to Produce?
We calculate the production cost per unit by dividing total production costs by the total number of units produced.
```{r}

# Average profit per unit (Assuming selling price is $500)
selling_price_per_unit <- 500
avg_profit_per_unit <- selling_price_per_unit - cost_per_unit
avg_profit_per_unit
```


```
How Much Profit Should Timeless Transport Models Expect to Make in the Second Half of 2005?
```

```{r}
# Calculate total expected profit for Jun-Dec 2005
total_forecasted_units <- sum(forecast_classic$mean) # Sum of forecasted units for Classic Cars
total_expected_profit <- total_forecasted_units * avg_profit_per_unit
total_expected_profit

```

Conclusion
Forecasted Sales: The sales for Classic Cars and other products have been forecasted for the second half of 2005 using the ARIMA model.
Production Cost: The production cost per unit was calculated to be approximately $100.91.
Average Profit: The average profit per unit is $399.09.
Expected Profit: Based on forecasted sales and average profit, Timeless Transport Models is expected to generate approximately $678,504.90 in profit for the second half of 200

```{r}
```


